\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2018-06-13 write_up/mixfishsim_ecolmod.tex
Manuscript revisions

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}
%% Numbered without titles
%\bibliographystyle{model1a-num-names}
%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}
%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
%% APA style
%%\bibliographystyle{model5-names}\biboptions{authoryear}
%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}
%% `Elsevier LaTeX' style
%%\bibliographystyle{elsarticle-num}
%%\bibliographystyle{elsarticle-harv}

%%%%%%%%%%%%%%%%%%%%%%%

@@

%%\title{A simulation framework for exploring spatiotemporal dynamics in mixed
%%	fisheries}

@@

\begin{abstract}
%\textcolor{gray}{[Guidance: A concise and factual abstract is required. The abstract 
%should state briefly the purpose of the research, the principal results and
%major conclusions. An abstract is often pnresented separately from the article,
%so it must be able to stand alone. For this reason, References should be
%avoided, but if essential, then citep the author(s) and year(s). Also,
%non-standard or uncommon abbreviations should be avoided, but if essential they
%must be defined at their first mention in the abstract itself.  Graphical
%abstract: Although a graphical abstract is optional, its use is encouraged as
%it draws more attention to the online article. The graphical abstract should
%summarize the contents of the article in a concise, pictorial form designed to
%capture the attention of a wide readership. Graphical abstracts should be
%submitted as a separate file in the online submission system. Image size:
%Please provide an image with a minimum of 531 X 1328 pixels (h X w) or
%proportionally more. The image should be readable at a size of 5 x 13 cm using
%a regular screen resolution of 96 dpi.  Preferred file types: TIFF, EPS, PDF or
%MS Office files.] \\}
%
\replaced[id = JJ]{Most fisheries}{Fishing} exploit\deleted[id=JJ]{s} spatially
and temporally hetergenous fish populations, using species-unselective gear
that can result in unintended, unwanted catch of low quota or protected
species. Reducing these unwanted catches is crucial for biological and economic
sustainability of `mixed fisheries' and implementation of an ecosystem approach
to fishing. \\

To implement effective spatial measures to reduce discards a good understanding
of spatiotemporal fishery dynamics is required. However, traditional
scientific advice is limited by a lack of highly resolved knowledge of
population distribution, movement and how fishers interact with different fish
populations. This reflects that data on fish location at high temporal and
spatial resolutions is expensive and difficult to collect and therefore proxies
inferred from either scientific surveys or commercial catch data are often used
to model distributions, often with limited spatial and temporal resolution. \\

To understand how resolution impacts mixed fisheries inference, we develop a
highly resolved spatiotemporal simulation model incorporating: i)
delay-difference population dynamics, ii) population movement using Gaussian
Random Fields to simulate patchy, hetergenously distributed populations, and
iii) fishery dynamics for multiple fleet characteristics based on targetting
via correlated random walk movement and learned behaviour. \\

@@

%\textcolor{gray}{[Guidance:: State the objectives of the work and provide an 
%adequate background, avoiding a detailed literature survey or a summary of the
%results.] \\}
%
Fishers exploit fish populations that are heterogenously distributed in space
and time with verying knowledge of species distributions using
species-unselective fishing gear. Fisheries that catch an assemblage of
species, known as mixed fisheries \replaced[id=JJ]{may}{, when managed by
	single-species quotas can end up} discard\deleted[id=JJ]{ing} overquota
catch \added[id=JJ]{when managed by single species quotas,} leading to
overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{This discarding of fish in
	excess of quota hampers the ability to limit fishing mortality to
	within sustainable limits \citep{Alverson1994, Crowder1998,
		Rijnsdorp2007}.} Reducing discarding is crucial to ensure
biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to
	fisheries}\replaced[id=PD]{and}{. As such} there is increasing interest
in technical solutions such as gear and spatial closures as ways of avoiding
discard\replaced[id=JJ]{ing of fish \citep{Kennelly2002, Catchpole2008,
		Bellido2011}}{s}.  \\

Use of spatial management as a tool has been proposed as a method to reduce
discards \added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}. However, its
implementation is hampered by lack of knowledge of fish and fishery
spatiotemporal dynamics and understanding of the scale at which processes are
important for management. Understanding the correct scale for spatial
management is crucial in order to implement measures at a resolution that
ensures effective management \citep{Dunn2016} while minimising economic impact.
For example, a scale that promotes species avoidance for vulnerable or low
quota species while allowing continuance of sustainable fisheries for available
quota species.\\

Ensuring measures are implemented at an appropriate scale has been a challenge
in the past that has led to ineffectual measures with unitended consequences
such as limited impact towards the management objective or increased benthic
impact on previously unexploited areas (e.g. the cod closure in the North
Sea \citep{Rijnsdorp2001,Dinmore2003}). Since then more refined spatial
information has become available through the combination of logbook and Vessel
Monitoring System (VMS) data \citep{Lee2010, Bastardie2010, Gerritsen2012,
	Mateo2016} and more real-time spatial management has been possible 
	\citep[e.g.][]{Holmes2011}. Such information is, however, patchy and
derived from an inherently biased sampling programme (i.e. targeted fishing).
Further, fishers generally only recorded landings (not catch) on a daily
basis\todo{\added[id=JJ]{Its not clear what the problem is: landings
		collected on daily basis or landings recorded rather than catch}}.
This leads to questions about the validity of inference that can be drawn from
landings data assigned to VMS activity pings\todo{\added[id=JJ]{This comes as a
	surprise: I thought this was going to be about discards}}.\\

In order to understand challenges that face VMS-linked landings to draw
inference on the underlying population structure we develop a simulation model
where population dynamics are highly-resolved in space and time and are known
rather than inferred from sampling or commercial catches. Population movement
is driven by a random (diffusive) and directed (advective) process and we
incorporate characterisation of a number of different fisheries exploiting four
fish populations with different spatial and population demographics.\\

Using our model we simulate 20 years of exploitation of the fish populations
and use the results from the fishing model to draw inference on the underlying
population structures.  We compare this inference to: i) a stratified
fixed-site sampling survey design commonly used for fisheries monitoring
purposes, otherwise know as a fisheries-independent
	survey, and ii) the underlying population structures input to the
simulation.\\

We simulate a fishery closure to protect one species based on the
fishery-dependent inferred distributions at a spatial and temporal scale
typical in fisheries management, and assess a theoretical "benefit" to the
population, and effect on the other three populations\todo{\added[id=JJ]{If
		the paper has two goals this should be clear from the start,
		but may be better over 2 MSs}}. Further, we extend our analysis
to a range of spatial and temporal scales to assess the impact of these
processes on the success of the management measure.\\

@@

%\textcolor{gray}{[Guidance: Provide sufficient details to allow the work to be
%reproduced by an independent researcher. Methods that are already published
%should be summarized, and indicated by a reference.  If quoting directly from a
%previously published method, use quotation marks and also citep the source. Any
%modifications to existing methods should also be described.] \\ }
%
\replaced[id=PD]{A}{We developed and implemented a simulation model with a}
modular event-based \replaced{simulation model was developed with}{approach,
	where} \added[id=PD]{sub-}modules \deleted[id=PD]{are} implemented on
independent time-scales appropriate to capture the characteristic of the
\replaced[id=PD]{different processes}{process modelled} (Figure 1).
\added[id=PD]{The following sub-modules were included to capture the full
	system: 1) Population dynamics, 2) Recruitment dynamics, 3) Population
	movement, 4) fishery dynamics.}\\

\deleted[id=PD]{The fishing model operated on a tow-by-tow basis, while}
\replaced[id=PD]{P}{p}opulation dynamics (fishing and natural mortality,
growth) operate on a daily time-step\replaced[id=PD]{, while p}{. P}opulation
movement occurs on a weekly time-step\replaced[id=PD]{. R}{, while r}ecruitment
\replaced[id=PD]{takes place}{occurs} periodcally each year for a set time
\replaced[id=PD]{duration}{period (e.g. 3 weeks)} at at specified point
individual to a species\deleted[id=PD]{.}\added[id=PD]{, while the fishing
	module operates on a tow-by-tow basis (multiple events a day)}. The
simulation framework is implemented in the statistical software package R
\citep{RCoreTeam2017} \replaced[id=PD]{and}{;} available as an R package from
the authors github (\url{www.github.com/pdolder/MixFishSim}).\\

@@

%\textcolor{gray}{[Guidance: A \underline{Theory} section should extend, not
%repeat, the background to the article already dealt with in the	Introduction
%and lay the foundation for further work.  In contrast, a
%\underline{Calculation} section represents a practical development from a
%theoretical basis.] \\}
%
\subsection{Population parameterisation}

@@

%\textcolor{gray}{[Guidance: Results should be clear and concise.]}

@@

%\textcolor{gray}{[Guidance: This should explore the significance of the results 
%of the work, not repeat them. A combined Results and Discussion section is
%often appropriate.  Avoid extensive citations and discussion of published
%literature.]}
%
\section{Conclusions}

%\textcolor{gray}{[Guidance: The main conclusions of the study may be presented 
%in a short Conclusions section, which may stand alone or form a subsection of a
%Discussion or Results and Discussion section.]}
%
\section*{Appendices}

%\textcolor{gray}{[Guidance: If there is more than one appendix, they should 
%be identified as A, B, etc.  Formulae and equations in appendices should be
%given separate numbering: Eq.  (A.1), Eq. (A.2), etc.; in a subsequent
%appendix, Eq. (B.1) and so on.  Similarly for tables and figures: Table A.1;
%Fig. A.1, etc.]}
%

>>>>

%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\begin{abstract}

\replaced[id = JJ]{Most fisheries}{Fishing} exploit\deleted[id=JJ]{s} spatially
and temporally heterogeneous fish populations, using species-unselective gear
that can result in unintended, unwanted catch of low quota or protected
species. Reducing these unwanted catches is crucial for biological and economic
sustainability of `mixed fisheries' and implementation of an ecosystem approach
to fishing. \\

\replaced[id = PD]{If fisheries are to avoid unwanted catch}{To implement
	effective spatial measures to reduce discards} a good understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distribution, movement and how fishers interact with different fish
populations. This reflects that data on fish location at high temporal and
spatial resolutions is expensive and difficult to collect and therefore proxies
inferred from either scientific surveys or commercial catch data are often used
to model distributions, often with limited spatial and temporal resolution. \\

To understand how resolution impacts mixed fisheries inference, we develop a
highly resolved spatiotemporal simulation model incorporating: i)
delay-difference population dynamics, ii) population movement using Gaussian
Random Fields to simulate patchy, heterogeneously distributed populations, and
iii) fishery dynamics for multiple fleet characteristics based on targeting
via correlated random walk movement and learned behaviour. \\

@@

Fishers exploit fish populations that are heterogeneously distributed in space
and time with varying knowledge of species distributions using
species-unselective fishing gear. \added[id=PD]{In doing so} fisheries
\deleted[id=PD]{that} catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} overquota catch \added[id=JJ]{when managed by
	single species quotas,} leading to overexploitation of fish populations
\added[id=JJ]{\citep{Ulrich2011a, Batsleer2015}}.  \added[id=JJ]{This
	discarding of fish in excess of quota hampers the ability to limit
	fishing mortality to within sustainable limits \citep{Alverson1994,
	Crowder1998, Rijnsdorp2007}} \deleted[id=PD]{; reducing
discarding is crucial} to ensure biological and
economic sustainability of fisheries\deleted[id=JJ]{ and implementation of an
	ecosystem approach to fisheries}\replaced[id=PD]{and}{. As such} there
is increasing interest in technical solutions such as gear and spatial closures
as ways of \replaced[id=PD]{reducing unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}}\citep{Kennelly2002,
		Catchpole2008, Bellido2011}.  \\

\replaced[id=PD]{Changes to spatial fishing patterns have}{Use of spatial
	management as a tools} been proposed as a method to reduce discards
\added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}. However,
\deleted[id=PD]{its} implementation is hampered by lack of knowledge of fish
and fishery spatiotemporal dynamics and understanding of the scale at which
processes are important for management. Understanding the correct scale for
spatial management is crucial in order to implement measures at a resolution
that ensures effective management \citep{Dunn2016} while minimising economic
impact.  For example, a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} an
appropriate scale has been a challenge in the past that has led to ineffectual
measures with unintended consequences such as limited impact towards the
management objective or increased benthic impact on previously unexploited
areas (e.g. the cod closure in the North Sea
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatial information has \added[id=PD]{since} become available through
the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}.  Such
information is, however, patchy and derived from an inherently biased sampling
programme (i.e. targeted fishing). \deleted[id=PD]{ Further, fishers generally only recorded
landings (not catch) on a daily basis\todo{\added[id=JJ]{Its not clear what the
		problem is: landings collected on daily basis or landings
		recorded rather than catch}}.  This leads to questions about
the validity of inference that can be drawn from landings data assigned to VMS
activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I thought this was
		going to be about discards}\added[id=PD]{Agree, have removed
		this to avoid confusion}}\\

In order to understand \replaced[id=PD]{the consequences of using}{challenges
	that face} VMS-linked landings to draw inference on the underlying
population structure we develop a simulation model where population dynamics
are highly-resolved in space and time\replaced[id=PD]{. Being}{ and are}
known \added[id=PD]{directly} rather than inferred from sampling or commercial
catches \added[id=PD]{, we can use the population model to evaluate how inference from
	fisheries-dependent and fisheries independent sampling relates to the
	real population structure}. \replaced[id=PD]{In our model system
	p}{P}opulation movement is driven by random (diffusive) and directed
(advective) processes and we incorporate characterisation of a number of
different fisheries \added[id=PD]{dynamics} exploiting four fish populations
with different spatial and population demographics.\\

Using our model we simulate \replaced[id=PD]{20}{40} years of exploitation of
the fish populations. \replaced[id=PD]{We}{ and} use the
results\deleted[id=PD]{ from the fishing model:}

\begin{enumerate}
	\item to understand how sampling-derived data reflects the underlying
		population structures. We compare at different spatial and
		temporal aggregations of data the real population to: 
		\begin{enumerate}
			\item the inferred population from a stratified fixed-site sampling survey design
				commonly used for fisheries monitoring purposes, otherwise know as a
				fisheries-independent survey,
			\item the inferred population from our fishery-dependent model which includes
				fishery-induced sampling dynamics.\\
		\end{enumerate}
	\item to understand the impact of data aggregation and source on
		spatial fisheries management measures we simulate a fishery closure to
		protect a species based on different spatial and temporal data
		aggregations:
		\begin{enumerate}
			\item as if the real spatial population structure were
				known,
			\item the fishery-independent inferred population
				structure
			\item the fishery-dependent inferred population
				structure
		\end{enumerate}
		We evaluate the theoretical "benefit" to the population of the
		closure(s), and effect on the other three
		populations\todo{\added[id=JJ]{If the paper has two goals this
				should be clear from the start, but may be
				better over 2 MSs}\added[id=PD]{I would like to
				keep both parts, but have made clearer in how
				its set out. The closure scenarios form
				validation of the data aggregation, rather than
				effectiveness of the closures themselves - so
				its a continuation of the same question in my
				eyes}}\deleted[id=PD]{.  Further, we extend our
			analysis to a range of spatial and temporal scales to
			assess the impact of these processes on the success of
			the management measure.}\\
\end{enumerate}

@@

\replaced[id=PD]{A}{We developed and implemented a simulation model with a}
modular event-based \replaced{simulation model was developed with}{approach,
	where} \added[id=PD]{sub-}modules \deleted[id=PD]{are} implemented on
independent time-scales appropriate to capture the characteristic of the
\replaced[id=PD]{different processes}{process modelled} (Figure 1).
\added[id=PD]{The following sub-modules were included to capture the full
	system: 1) Population dynamics, 2) Recruitment dynamics, 3) Population
	movement, 4) fishery dynamics.}\\

\deleted[id=PD]{The fishing model operated on a tow-by-tow basis, while}
\replaced[id=PD]{P}{p}opulation dynamics (fishing and natural mortality,
growth) operate on a daily time-step\replaced[id=PD]{, while p}{. P}opulation
movement occurs on a weekly time-step\replaced[id=PD]{. R}{, while r}ecruitment
\replaced[id=PD]{takes place}{occurs} periodcally each year for a set time
\replaced[id=PD]{duration}{period (e.g. 3 weeks)} at at specified point
individual to a species\deleted[id=PD]{.}\added[id=PD]{, while the fishing
	module operates on a tow-by-tow basis (i.e. multiple events a day)}.
The simulation framework is implemented in the statistical software package R
\citep{RCoreTeam2017} \replaced[id=PD]{and}{;} available as an R package from
the authors github (\url{www.github.com/pdolder/MixFishSim}).\\

@@

\subsection{Population parameterisation}

@@

@@

\section{Conclusions}

\section*{Appendices}
